Global land characterisation using land cover fractions at 100 m resolution

### 2021-02-18 article/article.tex
Replace UTF-8 characters

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{microtype}
\usepackage{amsmath}
%\usepackage{indentfirst}
\usepackage{booktabs} % For \toprule, etc.
\usepackage{textcomp} % for degree sign
%\usepackage[sort&compress,sectionbib]{natbib} % already in elsarticle
\usepackage{graphicx}
\usepackage{adjustbox} % For resizing tables
\usepackage{array} % For mid-alignment (m{xcm}) in tables
\usepackage{rotating} % Sideways figures
%\bibliographystyle{apalike} % elsarticle defined at the top

@@

Results showed that the random forest regression model reached the lowest RMSE of 17.3\%. Lowest MAE (7.9\%) and highest overall accuracy (72±2\%) was achieved using random forest with our proposed three-model approach and median vote.

@@

\begin{table}
\centering
%\adjustbox{max width=0.5\textwidth}{
\begin{tabular}{lllllll}
\toprule
\textbf{Model} & \textbf{\ac{RMSE} (\%)} & \textbf{\ac{MAE} (\%)} & $\mathbf{R^2_{NSE}}$ & $\mathbf{R^2_{OLS}}$ (slope/int) & \textbf{\ac{OA} (\%)} & \textbf{Kappa} \\
\midrule
\Gls{RF} regression
& \textbf{17.3}  & 9.4   & \textbf{0.66}  & \textbf{0.67} (1.09/-1.22) & 67±4  & 0.57±0.05 \\
\ensuremath{''} two-step
& 19.9  & 8.2   & 0.56  & 0.60 (0.78/3.48) & 71±2  & 0.62±0.02 \\
\ensuremath{''} three-step
& 19.4  & 8.4   & 0.58  & 0.60 (0.84/2.34) & 71±3  & 0.62±0.04 \\
\ensuremath{''} median vote
& 20.7  & \textbf{7.9}   & 0.52  & 0.60 (0.74/3.80) & 71±1  & 0.63±0.02 \\
\ensuremath{''} \ensuremath{''} two-step
& 20.0  & 8.1   & 0.54  & 0.60 (0.77/3.67) & \textbf{72±1}  & 0.63±0.02 \\
\ensuremath{''} \ensuremath{''} three-step
& 20.2  & \textbf{7.9}   & 0.54  & 0.60 (0.77/3.34) & \textbf{72±2}  & \textbf{0.64±0.02} \\
\bottomrule
\end{tabular}
%}
\caption{Accuracy statistics of multi-step models. Best performing statistics are highlighted. ``slope'' refers to the OLS-estimated slope, ``int'' refers to the OLS-estimated intercept.}
\label{tab-accuracy}
\end{table}

@@

We investigated ways to tackle the issue of accurately predicting the extreme fraction values of 0\% and 100\% by proposing a hierarchical multi-step approach combining classification and regression models.
This approach was applied to an \gls{RF} regression model, which our tests showed to have the highest accuracy (RMSE: 17.3\%) for land cover fraction mapping compared to other regression algorithms.
We also combined this approach with \gls{RF} median voting.
The combined \gls{RF} median three-step approach obtained the best results for \gls{MAE} (7.9\%) and \gls{SCM} \gls{OA} (72±2\%).
The proposed approach results in predictions of the most likely pure land cover class, when the class is uncertain.
This is in contrast to predicting a mix of classes when a standard one-step model is used, and therefore is useful for users who are more interested in the most likely class, rather than class probabilities.
Based on this model, we created a demonstration map showing the global distribution of land cover in separate land cover fraction layers.
Remote sensing features were the most important for model accuracy, although all other types of features (climate, soil, terrain) also contributed significantly for some classes and thus could not be omitted without negatively affecting the overall model accuracy.

@@

\begin{table}
\centering
%\adjustbox{max width=0.5\textwidth}{
\begin{tabular}{lllllll}
\toprule
\textbf{Model} & \textbf{\ac{RMSE} (\%)} & \textbf{\ac{MAE} (\%)} & $\mathbf{R^2_{NSE}}$ & $\mathbf{R^2_{OLS}}$ (slope/int) & \textbf{\ac{OA} (\%)} & \textbf{Kappa} \\
\midrule
Equal proportions
& 29.9  & 21.4  & 0     & — & 26±5  & 0.13±0.07 \\
\Glsentryshort{FNC}
& 24.4  & 13.5  & 0.33  & 0.35 (0.80/2.89) & 53±4  & 0.42±0.06 \\
\Glsentryshort{GLM}, \Glsentryshort{PLS}, Lasso
& 21.6  & 12.7  & 0.48  & 0.49 (1.10/-1.42) & 56±4  & 0.43±0.05 \\
\Glsentryshort{MLR}
& 21.6  & 12.1  & 0.48  & 0.48 (0.96/0.62) & 58±4  & 0.46±0.06 \\
\Glsentryshort{MLP} \glsentryshortpl{NN}
& 22.7  & 9.2   & 0.43  & 0.52 (0.70/4.25) & 68±1  & 0.57±0.02 \\
\Glsentryshort{SVM} regression
& 20.7  & 8.9   & 0.52  & 0.56 (0.79/3.02) & 69±2  & 0.58±0.03 \\
Cubist regression
& 18.1  & \textbf{8.1}   & 0.63  & 0.65 (0.88/1.77) & \textbf{72±2}  & \textbf{0.63±0.03} \\
\Glsentryshort{RF} regression
& \textbf{17.3}  & 9.4   & \textbf{0.66}  & \textbf{0.67} (1.09/-1.22) & 67±4  & 0.57±0.05 \\
\ensuremath{''}
only RS features
& 18.4  & 10.3  & 0.62  & 0.64 (1.09/-1.25) & 64±4  & 0.54±0.05 \\
Cubist + RF ensemble
& 17.7  & 8.6   & 0.65  & 0.65 (0.95/0.73) & 70±3 & 0.61±0.04 \\
\bottomrule
\end{tabular}
%}
\caption{Accuracy statistics of the tested regression models. Best performing statistics are highlighted. ``slope'' refers to the OLS-estimated slope, ``int'' refers to the OLS-estimated intercept. ``Only RS features'' stands for a model trained only with \glspl{VI} and temporal metrics.}
\label{tab-accuracy-linear}
\end{table}

>>>>

%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{microtype}
\usepackage{amsmath}
%\usepackage{indentfirst}
\usepackage{booktabs} % For \toprule, etc.
\usepackage{textcomp} % for degree sign
\usepackage[separate-uncertainty=true, multi-part-units=repeat, detect-weight=true]{siunitx}
%\usepackage[sort&compress,sectionbib]{natbib} % already in elsarticle
\usepackage{graphicx}
\usepackage{adjustbox} % For resizing tables
\usepackage{array} % For mid-alignment (m{xcm}) in tables
\usepackage{rotating} % Sideways figures
%\bibliographystyle{apalike} % elsarticle defined at the top

@@

Results showed that the random forest regression model reached the lowest RMSE of 17.3\%. Lowest MAE (7.9\%) and highest overall accuracy (\SI{72 \pm 2}{\percent}) was achieved using random forest with our proposed three-model approach and median vote.

@@

\begin{table}
\centering
%\adjustbox{max width=0.5\textwidth}{
\begin{tabular}{lllllll}
\toprule
\textbf{Model} & \textbf{\ac{RMSE} (\%)} & \textbf{\ac{MAE} (\%)} & $\mathbf{R^2_{NSE}}$ & $\mathbf{R^2_{OLS}}$ (slope/int) & \textbf{\ac{OA} (\%)} & \textbf{Kappa} \\
\midrule
\Gls{RF} regression
& \textbf{17.3}  & 9.4   & \textbf{0.66}  & \textbf{0.67} (1.09/-1.22) & \num{67\pm4}  & \num{0.57\pm0.05} \\
\ensuremath{''} two-step
& 19.9  & 8.2   & 0.56  & 0.60 (0.78/3.48) & \num{71\pm2}  & \num{0.62\pm0.02} \\
\ensuremath{''} three-step
& 19.4  & 8.4   & 0.58  & 0.60 (0.84/2.34) & \num{71\pm3}  & \num{0.62\pm0.04} \\
\ensuremath{''} median vote
& 20.7  & \textbf{7.9}   & 0.52  & 0.60 (0.74/3.80) & \num{71\pm1}  & \num{0.63\pm0.02} \\
\ensuremath{''} \ensuremath{''} two-step
& 20.0  & 8.1   & 0.54  & 0.60 (0.77/3.67) & \textbf{\num{72\pm1}}  & \num{0.63\pm0.02} \\
\ensuremath{''} \ensuremath{''} three-step
& 20.2  & \textbf{7.9}   & 0.54  & 0.60 (0.77/3.34) & \textbf{\num{72\pm2}}  & \textbf{\num{0.64\pm0.02}} \\
\bottomrule
\end{tabular}
%}
\caption{Accuracy statistics of multi-step models. Best performing statistics are highlighted. ``slope'' refers to the OLS-estimated slope, ``int'' refers to the OLS-estimated intercept.}
\label{tab-accuracy}
\end{table}

@@

We investigated ways to tackle the issue of accurately predicting the extreme fraction values of 0\% and 100\% by proposing a hierarchical multi-step approach combining classification and regression models.
This approach was applied to an \gls{RF} regression model, which our tests showed to have the highest accuracy (RMSE: 17.3\%) for land cover fraction mapping compared to other regression algorithms.
We also combined this approach with \gls{RF} median voting.
The combined \gls{RF} median three-step approach obtained the best results for \gls{MAE} (7.9\%) and \gls{SCM} \gls{OA} (\SI{72\pm2}{\percent}).
The proposed approach results in predictions of the most likely pure land cover class, when the class is uncertain.
This is in contrast to predicting a mix of classes when a standard one-step model is used, and therefore is useful for users who are more interested in the most likely class, rather than class probabilities.
Based on this model, we created a demonstration map showing the global distribution of land cover in separate land cover fraction layers.
Remote sensing features were the most important for model accuracy, although all other types of features (climate, soil, terrain) also contributed significantly for some classes and thus could not be omitted without negatively affecting the overall model accuracy.

@@

\begin{table}
\centering
%\adjustbox{max width=0.5\textwidth}{
\begin{tabular}{lllllll}
\toprule
\textbf{Model} & \textbf{\ac{RMSE} (\%)} & \textbf{\ac{MAE} (\%)} & $\mathbf{R^2_{NSE}}$ & $\mathbf{R^2_{OLS}}$ (slope/int) & \textbf{\ac{OA} (\%)} & \textbf{Kappa} \\
\midrule
Equal proportions
& 29.9  & 21.4  & 0     & — & \num{26\pm5}  & \num{0.13\pm0.07} \\
\Glsentryshort{FNC}
& 24.4  & 13.5  & 0.33  & 0.35 (0.80/2.89) & \num{53\pm4}  & \num{0.42\pm0.06} \\
\Glsentryshort{GLM}, \Glsentryshort{PLS}, Lasso
& 21.6  & 12.7  & 0.48  & 0.49 (1.10/-1.42) & \num{56\pm4}  & \num{0.43\pm0.05} \\
\Glsentryshort{MLR}
& 21.6  & 12.1  & 0.48  & 0.48 (0.96/0.62) & \num{58\pm4}  & \num{0.46\pm0.06} \\
\Glsentryshort{MLP} \glsentryshortpl{NN}
& 22.7  & 9.2   & 0.43  & 0.52 (0.70/4.25) & \num{68\pm1}  & \num{0.57\pm0.02} \\
\Glsentryshort{SVM} regression
& 20.7  & 8.9   & 0.52  & 0.56 (0.79/3.02) & \num{69\pm2}  & \num{0.58\pm0.03} \\
Cubist regression
& 18.1  & \textbf{8.1}   & 0.63  & 0.65 (0.88/1.77) & \textbf{\num{72\pm2}}  & \textbf{\num{0.63\pm0.03}} \\
\Glsentryshort{RF} regression
& \textbf{17.3}  & 9.4   & \textbf{0.66}  & \textbf{0.67} (1.09/-1.22) & \num{67\pm4}  & \num{0.57\pm0.05} \\
\ensuremath{''}
only RS features
& 18.4  & 10.3  & 0.62  & 0.64 (1.09/-1.25) & \num{64\pm4}  & \num{0.54\pm0.05} \\
Cubist + RF ensemble
& 17.7  & 8.6   & 0.65  & 0.65 (0.95/0.73) & \num{70\pm3} & \num{0.61\pm0.04} \\
\bottomrule
\end{tabular}
%}
\caption{Accuracy statistics of the tested regression models. Best performing statistics are highlighted. ``slope'' refers to the OLS-estimated slope, ``int'' refers to the OLS-estimated intercept. ``Only RS features'' stands for a model trained only with \glspl{VI} and temporal metrics.}
\label{tab-accuracy-linear}
\end{table}
